---
title: "Data607_Project_1"
author: "Femi Adesanya"
date: "February 24, 2018"
output: html_document
---
#### Project Summary

##### In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players: Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents For the first player, the information would be: Gary Hua, ON, 6.0, 1794, 1605


#### Data Import

##### Text file imported from local folder
```{r}
tdata <- read.csv(paste0("C:\\Users\\26291\\Documents\\Data_607\\tournamentinfo.txt"), header =F)
head(tdata)
```


```{r}
tail(tdata)
```


#### Data Preprocessing and Wrangling

##### To remove the header in order to retreive information related to player and rating
```{r}
tdata <- tdata[-c(1:4),]
head(tdata)
```

```{r}
pInfo <- tdata[seq(1, length(tdata), 3)]
rInfo <- tdata[seq(2, length(tdata), 3)]
```


#### Data Extraction
```{r}
library(stringr)
pairNo <- as.integer(str_extract(pInfo, "\\d+"))
Name <- str_trim(str_extract(pInfo, "(\\w+\\s){2,3}"))
Region <- str_extract(rInfo, "\\w+")
Points <- as.numeric(str_extract(pInfo, "\\d+\\.\\d+"))
Rating <- as.integer(str_extract(str_extract(rInfo, "[^\\d]\\d{3,4}[^\\d]"), "\\d+"))
Opponents <- str_extract_all(str_extract_all(pInfo, "\\d+\\|"), "\\d+")
Won <- str_count(pInfo, "\\Q|W  \\E")
Loose <- str_count(pInfo, "\\Q|L  \\E")
Draw <- str_count(pInfo, "\\Q|D  \\E")
```


#### Average Opponent Calculation

###### to calculate average Rating, we add all the Opponents Pre Torunamanet Ratings and divide it by the total number of games played by the player.
```{r}
avgRating <- length(pInfo)
for (i in 1:length(pInfo)) { 
  avgRating[i] <- round(mean(Rating[as.numeric(unlist(Opponents[pairNo[i]]))]), digits = 0)
}
fdata <- data.frame(Name, Region, Points, Rating, avgRating);
```


#### Final DataSet 

##### with the five attributes
```{r}

colnames(fdata) <- c("Player's Name", "Player's State", "Total Number of Points", "Player's Pre-Rating", " Average Pre Chess Rating of Opponents")
library(DT)
datatable(fdata)
```


#### Create csv file of the final dataset
```{r}
write.csv(fdata, file = "chessdata.csv")

```

