---
title: "Data607_Project2_Cleaning and Descriptive Statistics"
author: "Femi Adesanya"
date: "March 14, 2018"
output: 
  html_document: 
    df_print: kable
    fig_caption: yes
    theme: united
    toc: yes
    toc_depth: 6
---

```{r setup, include=FALSE}
pack = c("tidyverse", "magrittr", "psych", "GGally", 
         "readxl", "corkscrew", "knitr", "markdown")
sapply(pack, function(x) suppressMessages(require(x, character.only = T)))
knitr::opts_chunk$set(echo = F, message = F, warning = F,
                      error = F, fig.align="center")
```


## Cleaning Movie Data


```{r}
funs_v1 = function(x) {
  class = ifelse(is.character(x), "Categorical", "Continuous")
  nlev = ifelse(is.character(x), length(unique(na.omit(x))), NA) 
  nmiss = sum(is.na(x))
  nvalid = sum(!is.na(x))
  return(cbind(class, nlev, nmiss, nvalid) %>% as.tibble())
} 
```


###  <font color="darkblue">  Raw data information </font>


```{r}
setwd("C:\\Users\\26291\\Documents\\Data_607")
data_v1 = 
  read_csv("movie_metadata.csv", progress = F) %>%
  distinct() 

data_v1 %>% 
  map_df(funs_v1, .id = "name") %>% 
  arrange(class, name) %>%
  set_colnames(c("Name", "Type", "No. Levels", "No. Missing", "No. Valid")) %>%
  kable(caption = "Movie data it contains 4998 rows and 28 columns")

data_v1$movie_imdb_link = NULL
data_v1$plot_keywords = NULL
```


### <font color="darkblue">  Categorical variables </font>


#### <font color="darkred">  Split movie *genres* variable </font>
 

```{r}
genres_df = 
  data_v1 %$%
  str_split(genres, "[|]") %>%
  map(as.tibble) %>% 
  map(data.table::transpose) %>% bind_rows() %>%
  set_colnames(paste("genre", 1:8, sep = "_"))

cbind(data_v1 %>% select(genres), genres_df) %>%
  as.tibble() %>%
  head(10) %>%
  kable(caption = "Spliting movie genres variable into many columns (Ten rows)")
```



#### <font color="darkred"> Reduce dimension of *content_rating* variables </font>


```{r} 
data_v1 %>% 
  count(content_rating, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n)) %>%
  select(-n) %>%
  kable(caption = "Variable content_rating befor binning", digits=3)

content_rating_df =
  transmute(data_v1,
    content_rating = 
      content_rating %>% as.character() %>%
      recode(`TV-Y`="TV > 14", `TV-Y7`="TV > 14", `TV-Y`="TV > 14", `TV-PG`="TV > 14",   
             `TV-G`="TV > 14", `TV-PG`="TV > 14", `TV-14`="TV > 14", 
             `TV-Y17` = "TV > 14", `TV-MA` = "TV > 14", `NC-17`="TV > 14",
             M = "G-M-X", X = "G-M-X", G = "G-M-X",
             Unrated = "Unrated", `Not Rated` = "Unrated",
             PG = "PG-13", `PG-13`="PG-13", GP = "PG-13", 
             Approved = "Passed", Passed = "Passed"))

content_rating_df %>% 
  count(content_rating, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n) %>% round(2)) %>%
  select(-n) %>%
  kable(caption = "Variable content_rating after binning", digits=3)
```


#### <font color="darkred"> Reduce dimension of *country* variables </font> 


```{r}
data_v1 %>% 
  count(country, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n)) %>%
  select(-n) %>%
  top_n(10, Frequency) %>%
  kable(caption = "Variable country befor binning (Top 10)", digits=3)

country_df =
  transmute(data_v1,
    country = 
      country %>% as.character() %>%
      recode(USA = "USA", UK = "UK", .default = "Others"))
          #France="France", Canada="Canada", Germany="Germany", Australia="Australia",

country_df %>% 
  count(country, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n) %>% round(2)) %>%
  select(-n) %>%
  kable(caption = "Variable country after binning", digits=3)
```


#### <font color="darkred"> Reduce dimension of *language* variables </font> 


```{r}
data_v1 %>% 
  count(language, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n)) %>%
  select(-n) %>%
  top_n(10, Frequency) %>%
  kable(caption = "Variable language befor binning (Top 10)", digits=3)

language_df =
  transmute(data_v1,
    language = 
      language %>% as.character() %>%
      recode(English = "English", .default = "Others"))

language_df %>% 
  count(language, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n)) %>%
  select(-n) %>%
  kable(caption = "Variable language after binning", digits=3)
```


#### <font color="darkred"> Vizualisation </font>


```{r}
data_v2 = 
  data_v1 %>%
  select(-genres, -language, -country, -content_rating) %>%
  cbind(content_rating_df, country_df, language_df, genres_df) %>%
  as.tibble()

funs_v2 = function(x, title = " ") {
  xx = barplot(x, border = "dark blue", col = "white", las=2, 
          cex.axis = .9, cex.names = 0.7, ylab = "Frequency", main = title, 
          ylim = c(0, ceiling(1.1*max(x))))
  text(x = xx, y=as.numeric(x), label=as.numeric(x), 
       pos=3, cex = .7, col = "darkblue")
}
```


##### <font color="purple"> Movie genres level </font> 


```{r, fig.width=12, fig.height=8}

par(mfrow = c(2,2))
data_v2$genre_1 %>% table() %>% sort() %>% funs_v2(title = "Genre level 1")
data_v2$genre_2 %>% table() %>% sort() %>% funs_v2(title = "Genre level 2")
data_v2$genre_3 %>% table() %>% sort() %>% funs_v2(title = "Genre level 3")
data_v2$genre_4 %>% table() %>% sort() %>% funs_v2(title = "Genre level 4")
data_v2$genre_5 %>% table() %>% sort() %>% funs_v2(title = "Genre level 5")
data_v2$genre_6 %>% table() %>% sort() %>% funs_v2(title = "Genre level 6")
data_v2$genre_7 %>% table() %>% sort() %>% funs_v2(title = "Genre level 7")
data_v2$genre_8 %>% table() %>% sort() %>% funs_v2(title = "Genre level 8")
par(mfrow = c(1,1))
```


##### <font color="purple"> Movie top actors and directors </font> 


```{r, fig.width=12, fig.height=8}
n = 20
par(mfrow = c(2,2))
data_v2$actor_1_name %>% table() %>% sort(decreasing = T) %>% .[1:n] %>%
  funs_v2(title = "Top 20 principal actors")
data_v2$actor_2_name %>% table() %>% sort(decreasing = T) %>% .[1:n] %>%
  funs_v2(title = "Top 20 2nd actors")
data_v2$actor_3_name %>% table() %>% sort(decreasing = T) %>% .[1:n] %>%
  funs_v2(title = "Top 20 3rd actors")
data_v2$director_name %>% table() %>% sort(decreasing = T) %>% .[1:n] %>%
  funs_v2(title = "Top 20 3rd director")
par(mfrow = c(1,1))
```


##### <font color="purple"> Movies other variables </font> 


```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,2))
data_v2$color %>% table() %>% sort(decreasing = T) %>%
  funs_v2(title = "Movies color")
data_v2$country %>% table() %>% sort(decreasing = T) %>%
  funs_v2(title = "Movies country")
data_v2$language %>% table() %>% sort(decreasing = T) %>%
  funs_v2(title = "Movies language")
data_v2$content_rating %>% table() %>% sort(decreasing = T) %>%
  funs_v2(title = "Movies content rating")
par(mfrow = c(1,1))
```


### <font color="darkblue">  Continuous variables </font>

#### <font color="darkred"> Descriptive statistics </font>

```{r}
continous = sapply(data_v2, is.character)
data_v3 = data_v2[,!continous]

describe(data_v3 %>% select(-title_year)) %>%
  mutate(name = rownames(.)) %>%
  select(name, mean, median, min, max) %>%
  arrange(name) %>%
  kable(caption = "Descriptive statistics", digits = 2)
```


```{r}
n=10
data_v3 %>% 
  count(title_year, sort = T) %>%
  mutate(Frequency = n,
         Proportion = n/sum(n)) %>%
  select(-n) %>%
  top_n(n, Frequency) %>%
  kable(caption = "Top 10 years", digits = 3)
```


```{r, fig.width=14, fig.height=5}
x = data_v3$title_year %>% table() %>% sort(decreasing = T) %>% .[1:n]
xx = barplot(x, border = "dark blue", col = "white", las=2, 
             cex.axis = .9, cex.names = 0.7, ylab = "No. movies", 
             ylim = c(0, ceiling(1.1*max(x))))
text(x = xx, y=as.numeric(x), label=as.numeric(x), 
     pos=3, cex = 0.7, col = "darkblue", crt=45)

```

<!---
#### <font color="darkred"> Vizualisation </font>

```{r}
funs_v3 = function(x, title) {
  par(mfrow=c(1,2))
  d = density(x, na.rm = T)
  plot(d, main=title, xlab = "")
  polygon(d, col="#a32020", border="darkblue")
  boxplot(x, col = "#a32020", border = "darkblue", main = title, ylab = "Values")
  par(mfrow=c(1,1))
}
```



```{r, fig.width=10, fig.height=5, eval=F}
data_v3$num_critic_for_reviews %>% funs_v3(title = "No. critic for reviews")
data_v3$duration %>% funs_v3(title = "Duration")
data_v3$actor_1_facebook_likes %>% funs_v3(title = "No. facebook likes 1st actor")
data_v3$actor_2_facebook_likes %>% funs_v3(title = "No. facebook likes 2nd actor")
data_v3$actor_3_facebook_likes %>% funs_v3(title = "No. facebook likes 3rd actor")
data_v3$director_facebook_likes %>% funs_v3(title = "No. facebook likes director")
data_v3$gross %>% funs_v3(title = "Gross")
data_v3$num_voted_users %>% funs_v3(title = "No. votes")
data_v3$cast_total_facebook_likes %>% funs_v3(title = "No. cast facebook likes")
data_v3$facenumber_in_poster %>% funs_v3(title = "No. face in poster")
data_v3$num_user_for_reviews %>% funs_v3(title = "No. reviews")
data_v3$budget %>% funs_v3(title = "Budget")
#data_v3$title_year %>% funs_v3(title = "")
data_v3$imdb_score %>% funs_v3(title = "IMBD score")
data_v3$aspect_ratio %>% funs_v3(title = "Aspect ratio")
data_v3$movie_facebook_likes %>% funs_v3(title = "No. movie facebook likes")
```
--->


